---
title: 'ANA 515 P assignment 1  '
author: "Chester Kayonga"
date: "2024-06-21"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "2020 NFL Predictions Analysis"
author: "Your Name"
date: "2024-06-21"
output: html_document
---

##A Description of the Data

The dataset I am working with contains game-by-game Elo ratings for the 2020 NFL season. This data IS from two sheets in an Excel workbook that we will combine into one. Our analysis aims to clean and explore the data, identify any errors, and produce visualizations to aid in understanding the data.

##Read the Data into R

```{r, include=FALSE}
# Load necessary libraries
library(readxl)
library(tidyverse)
```

```{r, include=FALSE}
# Read data from the Excel file
my_sheet1 <- read_excel("nfl_elo_latest (1).xlsx", sheet = "Sheet 1")
my_sheet2 <- read_excel("nfl_elo_latest (1).xlsx", sheet = "Sheet1")
```

##Combining Datafiles
```{r, include=FALSE}
# Combine the data from the two sheets
combined_data <- rbind(my_sheet1, my_sheet2)
```
##Cleaning the Data
##Data Inspection
```{r, echo=TRUE}
# Display the first few rows of the combined dataset
head(combined_data)

# Check for any missing values
missing_values <- sum(is.na(combined_data))
missing_values
```

##Handling Missing values

```{r, echo=TRUE}
# Identify columns with missing values
cols_with_na <- colnames(combined_data)[apply(combined_data, 2, anyNA)]
cols_with_na

# Impute missing values or remove rows/columns with missing values
# For simplicity, we will remove the rows with missing values
cleaned_data <- combined_data %>% drop_na()

# Verify that there are no missing values left
sum(is.na(cleaned_data))

```
##Taking care of Duplicates 
```{r echo=TRUE}
# Identify potential duplicate rows
duplicates <- combined_data[duplicated(combined_data), ]
duplicates

# Remove duplicate rows
cleaned_data <- cleaned_data %>% distinct()

# Display the first few rows of the cleaned dataset
head(cleaned_data)
```
##Ensuring 'elo_pre is Numeric

```{r include=FALSE}
# Convert the elo1_pre column to numeric
cleaned_data$elo1_pre <- as.numeric(cleaned_data$elo1_pre)

# Check for any NA values introduced during conversion
sum(is.na(cleaned_data$elo1_pre))
```
##Data visualization
##Histogram of Elo Ratings
```{r echo=TRUE}
# Create a histogram of Elo ratings (using 'elo1_pre' column)
ggplot(cleaned_data, aes(x = elo1_pre)) +
  geom_histogram(binwidth = 50, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Elo Ratings",
       x = "Elo Rating",
       y = "Frequency") +
  theme_minimal()
```

##Scatter Plot of Elo Ratings vs. Probability
```{r echo=TRUE}
# Create a scatter plot of Elo ratings vs. probability of winning (using 'elo1_pre' and 'elo_prob1' columns)
ggplot(cleaned_data, aes(x = elo1_pre, y = elo_prob1)) +
  geom_point(color = "blue", alpha = 0.5) +
  labs(title = "Scatter Plot of Elo Ratings vs. Probability of Winning",
       x = "Elo Rating",
       y = "Probability of Winning") +
  theme_minimal()
```

